---
title: "HW2 - Firm Size in Ethiopia"
output: pdf_document
date: "2024-03-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Loading libraries ---------------------------------
library(tidyverse)
library(gt)
library(haven)
library(sf)
library(glue)
library(modelsummary)
library(fixest)
library(fastDummies)
library(ggplot2)
library(ggalt)

data <- read_dta("middle_01_all_clean.dta")

size <-
  data |> 
  group_by(sample) |> 
  count()

```

##Question 1: -----


```{r}
#Histogram of firm employment
hist(data$emp, breaks = 1000, freq = FALSE, xlim = c(0,100), 
     main = "Distribution of Firm Employment in Ethiopia (Small & Medium Enterprises)",
     xlab = "Number of employees",
     ylab = "Share of Plants")  

# Fitting a density function to the data
density_fit <- density(data$emp) #what is this for

# Adding the line of best fit
lines(density_fit, col = "red", lwd = 2) #why is this code not working 
```

It would be misleading to only use Sample 1 as sample 1 focuses on the medium
and large sized enterprises and using those would skew the mean employment in
a firm to be much larger than it actually is. It would also not be representative
of true firm employment distribution in Ethiopia as Sample 2 makes up a considerable portion of the number of firms, and on a macro scale, the true employment distribution is skewed largely to the left, indicating more small-sized firms, and removing sample 1 would change this distribution drastically.

##Question 2: ------

```{r}
#histogram of firm sales
hist(data$sales, breaks = 50, freq = FALSE, 
     main = "Distribution of Sales in Ethiopia (Small & Medium Enterprises)",
     xlab = "Number of sales",
     ylab = "Share of Plants",)

# Fitting a density function to the data
density_fit2 <- density(data$sales, na.rm = TRUE)

# Adding the line of best fit
lines(density_fit2, col = "blue", lwd = 2)
```


Some of the possible reasons for the differences include: industry and sector
differences, where some small firms may be specialized in certain industries 
and may have achieved economies of scale, allowing them to produce a very high
number of sales with fewer employees. Furthermore, it's important to consider
the type of output produced by large and medium firms, which may have a lower
market price if it's mass produced compared to small firms, therefore, requiring
more sales to break even. 

## Question 3: ------

As can be seen by the regression line, there is a positive correlation between 
firm age and frim employment. One can take from this that older firms are ones
that survived market competition and thus were able to grow as they were the 
most productive. It could also be that it takes many years to be able to grow the employment of your firm substantially.
However, if you look at the outliers (here firms with employment greater than 1200), you can see that while there is a greater concentration of outliers at the older age mark, age is not a restriction of firms becoming very large. One explanation on the concentration of very large firms at the oldest age is that these could be the large state owned enterprises (SOEs) in Ehtiopia.

```{r}
outliers <- data[data$emp > 1200,]
ggplot(data = data, aes(x = age, y = emp)) +
  geom_count(col= "tomato3", show.legend = F) +
  geom_smooth(method = "lm", se = FALSE) +  # Add a linear regression line
  labs(title = "Relationship Between Firm Employment and Age", x = "Firm Age", y = "Firm Employment") +
  geom_encircle(aes(x = age, y = emp),
                data = outliers,
                color = "cadetblue")
```

## Question 4a: -----

$$
y_{i} = A_{i}k_{i}^{\alpha}l_{i}^{1-\alpha}
$$
$$
VMPK_{i} = \alpha A_{i}k_{i}^{\alpha-1}l_{i}^{1-\alpha}
$$
$$
VMPK_{i} = \alpha \frac{y_{i}}{k_{i}}
$$
## Question 4b: -----

```{r}
data <-
  data |> 
  mutate(vmpk = 0.3*(sales/fixedasset_end))

outlier <-
  data |> 
  filter(!is.na(vmpk)) |> 
  filter(row_number() <= n() * 0.95) |> 
  mutate(log_vmpk = log(vmpk))

summ <-
  outlier |> 
  summarise(mean = mean(log_vmpk),
            sd = sd(log_vmpk))

#cross check values of mean and sd

```

After removing outliers, the mean and standard of log(vmpk) in birr are -0.021 and 1.70 respectively.

## Question 4c: -----

```{r}
ggplot(outlier, aes(x= emp, y= log_vmpk)) +
    geom_point() +
    geom_smooth(method=lm, se=FALSE)

#bin scatter and should we cut at 100 employees?
#revisit interpretation

```

Based on the scatter plot, we can observe that small scale firms that have very few employees, tend to have a very noisy relationship with value of marginal product of capital. As firm sizes increase their value of marginal product of capital tends to stabilize around the regression line. The graph also indicates a negative relationship between the two. In the large firms, the value of marginal product of labor tends to be lower probably due to capital crunch. 

Using the production function above and based on this relationship, we can comment that the productivity in larger firms tend to be lower. This is because as the increase in sales is not proportional to increase in capital and points to the lower contribution of labor. 


