---
title: "HW2 - Firm Size in Ethiopia"
author: "Nikhilla B, Ezeriki Emetonjor, Alistair Pyle, Suman Padhi"
output: pdf_document
date: "2024-03-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Loading libraries ---------------------------------
library(tidyverse)
library(gt)
library(haven)
library(sf)
library(glue)
library(modelsummary)
library(fixest)
library(fastDummies)
library(ggplot2)
library(ggalt)
library(labelled)
library(binsreg)

data <- read_dta("middle_01_all_clean.dta")

size <-
  data |> 
  group_by(sample) |> 
  count()

```

## Question 1: -----


```{r Question 1, echo=FALSE, message=FALSE, warning=FALSE}
#Histogram of firm employment
hist(data$emp, breaks = 1000, freq = FALSE, xlim = c(0,100), 
     main = "Distribution of Firm Employment in Ethiopia (Small & Medium Enterprises)",
     xlab = "Number of employees",
     ylab = "Share of Plants")  

# Fitting a density function to the data
density_fit <- density(data$emp) #what is this for

# Adding the line of best fit
lines(density_fit, col = "red", lwd = 2) #why is this code not working 
```

It would be misleading to only use Sample 1 as sample 1 focuses on the medium
and large sized enterprises and using those would skew the mean employment in
a firm to be much larger than it actually is. It would also not be representative
of true firm employment distribution in Ethiopia as Sample 2 makes up a considerable portion of the number of firms, and on a macro scale, the true employment distribution is skewed largely to the left, indicating more small-sized firms, and removing sample 1 would change this distribution drastically.

## Question 2: ------

```{r echo=FALSE, message=FALSE, warning=FALSE}
#histogram of firm sales
hist(data$sales, breaks = 50, freq = FALSE, 
     main = "Distribution of Sales in Ethiopia (Small & Medium Enterprises)",
     xlab = "Number of sales",
     ylab = "Share of Plants",)

# Fitting a density function to the data
density_fit2 <- density(data$sales, na.rm = TRUE)

# Adding the line of best fit
lines(density_fit2, col = "blue", lwd = 2)
```


Some of the possible reasons for the differences include: industry and sector
differences, where some small firms may be specialized in certain industries 
and may have achieved economies of scale, allowing them to produce a very high
number of sales with fewer employees. Furthermore, it's important to consider
the type of output produced by large and medium firms, which may have a lower
market price if it's mass produced compared to small firms, therefore, requiring
more sales to break even. Lastly, the absolute value of small firms present in Ethiopia, making up a large share of employment, means that their sales would generate a large portion of sales.

## Question 3: ------

As can be seen by the regression line, there is a positive correlation between 
firm age and firm employment. One can take from this that older firms are ones
that survived market competition and thus were able to grow as they were the 
most productive. It could also be that it takes many years to be able to grow the employment of your firm substantially.
However, if you look at the outliers (here firms with employment greater than 1200), you can see that while there is a greater concentration of outliers at the older age mark, age is not a restriction of firms becoming very large. One explanation on the concentration of very large firms at the oldest age is that these could be the large state owned enterprises (SOEs) in Ehtiopia.

```{r echo=FALSE, message=FALSE, warning=FALSE}
#alternative code
graph <- binsreg(data$emp, data$age, line = c(1,1))
graph
```

## Question 4a: -----

$$
y_{i} = A_{i}k_{i}^{\alpha}l_{i}^{1-\alpha}
$$
$$
VMPK_{i} = \alpha A_{i}k_{i}^{\alpha-1}l_{i}^{1-\alpha}
$$
$$
VMPK_{i} = \alpha \frac{y_{i}}{k_{i}}
$$


## Question 4b: -----

```{r echo=FALSE, message=FALSE, warning=FALSE}
# filtering out na values of vmpk from the get go
data <-
  data |> 
  mutate(vmpk = 0.3*(sales/fixedasset_end)) |> 
  filter(!is.na(vmpk))

# calculating percentiles based on vmpk and filtering out those that
# that are more than 5%
data$quantile <- ecdf(data$vmpk)(data$vmpk)
mean(data$quantile)
#filtering out top 5%
outlier_2 <- data |>
  filter(quantile < 0.95) |>
  mutate(log_vmpk = log(vmpk))


summ <-
  outlier_2 |> 
  summarise(mean_vmpk = mean(log_vmpk),
            sd = sd(log_vmpk))

gt(summ)

#cross check values of mean and sd

```

After removing outliers, the mean and standard of log(vmpk) in birr are -0.211 and 1.45 respectively.

## Question 4c: -----

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(outlier_2, aes(x= emp, y= log_vmpk)) +
    geom_point() +
    geom_smooth(method=lm, se=FALSE)

#bin scatter and should we cut at 100 employees?
#revisit interpretation

```

Based on the scatter plot, we can observe that small scale firms that have very few employees, tend to have a very noisy relationship with value of marginal product of capital. As firm sizes increase their value of marginal product of capital tends to stabilize around the regression line. The graph also indicates a negative relationship between the two. In the large firms, the vmpk tends to be lower probably due to capital crunch for smaller firms. 

Using the production function above and based on this relationship, we can comment that the marginal product of capital in larger firms tend to be lower. This is because the increase in sales is not proportional to increase in capital and points to the lower contribution of labor. 


## Question 4d: -----
The results show us that firms that experienced working capital shortage had, on average, 24.4% higher VMPK than firms that did not experience such shortage. This does make sense: we have seen that smaller firms have higher values of VMPK and so this higher VMPK could be because they lack access to sufficient capital and experience capital crunches. So, possibly if smaller firms had better access to capital they could grow more and add more capital (which would decrease their VMPK).
```{r echo=FALSE, message=FALSE, warning=FALSE}
#The variables “prob_major1”, “prob_major2”, “prob_major3” report the three major problems faced by firms. Construct a variable that is equal to one if a firm reports “working capital shortage” (the code is 5) as any of the responses to these three variables; otherwise, it takes a value of zero. Run a regression that compares the log(VMPK) between firms that report working capital shortages and those that do not, while controlling for firm size. Provide an interpretation of the finding.

capital_analysis <- outlier_2 |>
  mutate(capital_short = case_when(prob_major1 == 5  ~ 1,
                                   prob_major2 == 5  ~ 1,
                                   prob_major3 == 5  ~ 1,
                                   TRUE ~ 0))

# running regressions
capital_lm <- lm(log_vmpk ~ capital_short + emp, data = capital_analysis)

# results
rename_map <- list(
  'capital_short' = 'Working Capital Shortage',
  'emp' = 'Employment')

models <- list("Log of VMPK" = capital_lm)

modelsummary(models,
             gof_map = "nobs",
             coef_map = rename_map)


```

## Question 4e: -----
From the results above, it seems that High Interest rate as a singular category has the strongest negative relationship with VMPK, in which a firm that has not solved its capital problem due to high interest rates sees about 161% less VMPK as compared to firms that cited insufficient loans as their reason. 

On the other hand, it seems that short loan of repayment actually has a positive relationship with VMPK, as firms that cited that reason actually see about 52.9% higher VMPK. This could be explained by the fact that having a shorter time to repay loans discourages firms from seeking capital, meaning they do not add capital into their firms, keeping VMPK higher.
```{r echo=FALSE, message=FALSE, warning=FALSE}
#Within Sample 2, there is a variable called “reason_not_solve_cap_prob” which takes on 6 possible values for why firms may be unable to solve their working capital problems. It’s only been answered by 167 firms, but let’s take the answers at face value and assume firms are telling the truth. Re-do the analysis in (d) comparing the log(VMPK) of firms controlling for firm size and the different responses to this question. What message emerges from the regression? (Hint, if you write “i.reason_not_solve_cap_prob”, Stata will automatically create the dummy variables for you and drop one (“insufficient loan”).

# getting the string representation of the reasons
capital_analysis$reason_not_solve_cap_string <- to_factor(capital_analysis$reason_not_solve_cap_prob)

# regression 2
# insufficient loan is also the reference
capital_notsolve_lm <- lm(log_vmpk ~ as.factor(reason_not_solve_cap_string) + emp + capital_short, data = capital_analysis)

# renaming covariates
rename_map2 <- list(
  'as.factor(reason_not_solve_cap_string)High collateral requirment' = 'High collateral requirment',
  'as.factor(reason_not_solve_cap_string)High interest rate' = 'High interest rate',
  'as.factor(reason_not_solve_cap_string)Short period of loan repayment' = 'Short period of loan repayment',
  'as.factor(reason_not_solve_cap_string)Inefficiency of financial institutions' = 'Inefficiency of financial institutions',
  'as.factor(reason_not_solve_cap_string)Others' = 'Others',
  'capital_short' = 'Working Capital Shortage',
  'emp' = 'Employment')

models <- list("Log of VMPK" = capital_notsolve_lm)

# results
modelsummary(models,
             gof_map = "nobs",
             coef_map = rename_map2,
             )

```

## Code:
```{r}
# Question 1
knitr::opts_chunk$set(echo = TRUE)

# Loading libraries ---------------------------------
library(tidyverse)
library(gt)
library(haven)
library(sf)
library(glue)
library(modelsummary)
library(fixest)
library(fastDummies)
library(ggplot2)
library(ggalt)
library(labelled)
library(binsreg)

data <- read_dta("middle_01_all_clean.dta")

size <-
  data |> 
  group_by(sample) |> 
  count()

# QUestion 2
#histogram of firm sales
hist(data$sales, breaks = 50, freq = FALSE, 
     main = "Distribution of Sales in Ethiopia (Small & Medium Enterprises)",
     xlab = "Number of sales",
     ylab = "Share of Plants",)

# Fitting a density function to the data
density_fit2 <- density(data$sales, na.rm = TRUE)

# Adding the line of best fit
lines(density_fit2, col = "blue", lwd = 2)

# Question 3
#alternative code
binsreg(data$emp, data$age, line = c(1,1))

# Question 4b
# filtering out na values of vmpk from the get go
data <-
  data |> 
  mutate(vmpk = 0.3*(sales/fixedasset_end)) |> 
  filter(!is.na(vmpk))

# calculating percentiles based on vmpk and filtering out those that
# that are more than 5%
data$quantile <- ecdf(data$vmpk)(data$vmpk)
mean(data$quantile)
#filtering out top 5%
outlier_2 <- data |>
  filter(quantile < 0.95) |>
  mutate(log_vmpk = log(vmpk))


summ <-
  outlier_2 |> 
  summarise(mean_vmpk = mean(log_vmpk),
            sd = sd(log_vmpk))

gt(summ)

# Question 4c
ggplot(outlier_2, aes(x= emp, y= log_vmpk)) +
    geom_point() +
    geom_smooth(method=lm, se=FALSE)

# Question 4d
capital_analysis <- outlier_2 |>
  mutate(capital_short = case_when(prob_major1 == 5  ~ 1,
                                   prob_major2 == 5  ~ 1,
                                   prob_major3 == 5  ~ 1,
                                   TRUE ~ 0))

# running regressions
capital_lm <- lm(log_vmpk ~ capital_short + emp, data = capital_analysis)

# results
rename_map <- list(
  'capital_short' = 'Working Capital Shortage',
  'emp' = 'Employment')

models <- list("Log of VMPK" = capital_lm)

modelsummary(models,
             gof_map = "nobs",
             coef_map = rename_map)

# Question 4e
# getting the string representation of the reasons
capital_analysis$reason_not_solve_cap_string <- to_factor(capital_analysis$reason_not_solve_cap_prob)

# regression 2
# insufficient loan is also the reference
capital_notsolve_lm <- lm(log_vmpk ~ as.factor(reason_not_solve_cap_string) + emp, data = capital_analysis)

# renaming covariates
rename_map2 <- list(
  'as.factor(reason_not_solve_cap_string)High collateral requirment' = 'High collateral requirment',
  'as.factor(reason_not_solve_cap_string)High interest rate' = 'High interest rate',
  'as.factor(reason_not_solve_cap_string)Short period of loan repayment' = 'Short period of loan repayment',
  'as.factor(reason_not_solve_cap_string)Inefficiency of financial institutions' = 'Inefficiency of financial institutions',
  'as.factor(reason_not_solve_cap_string)Others' = 'Others',
  'emp' = 'Employment')

models <- list("Log of VMPK" = capital_notsolve_lm)

# results
modelsummary(models,
             gof_map = "nobs",
             coef_map = rename_map2,
             )
```

